Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-24 article.tex
DE comments on the filter section

As the Figure~\ref{fig:original_curve} illustrates, the raw
fluorescence signal is very noisy due to light scattering artifacts that
ordinarily affects the quality of the recording~\citep{lichtman2011big}.
Accordingly, the first step of our pipeline is to smoothen the signal, using
one the following low-pass filter for filtering out high frequency noise :
\begin{align}
% symmetrical median filter
f_1(x^t_i) &= x^{t-1}_i + x^t_i + x^{t+1}_i \label{eq:symetric-median}, \\
% asymmetrical weighted median
f_2(x^t_i) &= 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i + x_i^t,
\label{eq:weighted-asymetric-median}
\end{align}
where $f_1$ effect is shown in Figure \ref{fig:lp_curve}.

Neurons communicates through short peaks, characterized by a high
frequency, while low frequencies of the signal corresponds to the slow
decay of fluorescence. The second step of our pipeline consists in
identifying neuron spikes by filtering out
low frequencies through a high-pass filter, resulting in the plot of
Figure \ref{fig:hp_curve}:
\begin{align}
g(x^{t}_{i}) &= x^{t}_i - x^{t-1}_i. \label{eq:high-pass-filter}
\end{align}

Once the low- and and high-pass filters have been applied, the remaining signal
corresponds to the finite backward difference of the fluorescence time series.
To further filter out small signal variations, mostly due to noise, the third
step consists in a  hard-threshold filter, yielding the Figure
\ref{fig:threshold_curve}:
\begin{align}
h(x^{t}_i) &= x^{t}_i \mathbb{1}(x^{t}_i \geq \tau) \text{ with } \tau > 0.
\end{align}

@@

% check dependency map => minimal d-map with distribution?
To a joint probability distribution $P_X$ corresponds a minimal dependency
graph. Each node of this graph is a random variable. Two random variables are
only connected if they are dependent. We consider that two neurons are
connected if  and only if,  there exists  in this dependency  graph an edge
that connects  directly the  two random  variables to  which they correspond.
This is equivalent to writing that two neurons $i$ and $j$ are  not  directly
connected  if and only if  $X_i$  and $X_j$  are  conditionally independent
with  respect to all  other variables. We remind  that two random  variables
$X_i$  and  $X_j$  are  said  to  be  conditionally independent to a (possibly
empty)  set of controlling random variables $Z$, denoted $X_i \perp X_j | Z$,
if the joint conditional probability distribution   $P_{X_i,  X_j   |   Z}$
factorizes  into   $P_{X_i|Z} P_{X_j|Z}$.  In particular,  when  $X_i$ and
$X_j$ are  conditionally independent  with respect  to all  the other
variables, denoted  $X_i \perp  X_j  | X^{-i,j}$,  then  the  observation  of
$X_i$  given  the knowledge  of  $X^{-i,j}$ provides  no  information  about
$X_j$,  and vice-versa.

>>>>

As the Figure~\ref{fig:original_curve} shows, the raw
fluorescence signal is very noisy due to light scattering artifacts that
ordinarily affects the quality of the recording~\citep{lichtman2011big}.
Accordingly, the first step of our pipeline is to smoothen the signal, using
one the following low-pass filter for filtering out high frequency noise :
\begin{align}
% symmetrical median filter
f_1(x^t_i) &= x^{t-1}_i + x^t_i + x^{t+1}_i \label{eq:symetric-median}, \\
% asymmetrical weighted median
f_2(x^t_i) &= 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i + x_i^t.
\label{eq:weighted-asymetric-median}
\end{align}
For the sake of illustration, the effect of the filter $f_1$ on the signal
is shown in Figure \ref{fig:lp_curve}.

Neurons communicate through short spikes, characterized by a high
frequency, while low frequencies of the signal correspond to the slow
decay of fluorescence. The second step of our pipeline consists in
a high-pass filter that transforms the time series into its backward
difference
\begin{align}
g(x^{t}_{i}) &= x^{t}_i - x^{t-1}_i. \label{eq:high-pass-filter},
\end{align}
resulting in the plot of Figure \ref{fig:hp_curve}:

To further filter out small signal variations, mostly due to noise, the third
step consists in a  hard-threshold filter
\begin{align}
h(x^{t}_i) &= x^{t}_i \mathbb{1}(x^{t}_i \geq \tau) \text{ with } \tau > 0,
\end{align}
yielding the Figure \ref{fig:threshold_curve}.

@@

% check dependency map => minimal d-map with distribution?
\textcolor{red}{This part is better, however there is still something
wrong with the theory.}
To a joint probability distribution $P_X$ corresponds a minimal dependency
graph. Each node of this graph is a random variable. Two random variables are
only connected if they are dependent. We consider that two neurons are
connected if  and only if,  there exists  in this dependency  graph an edge
that connects  directly the  two random  variables to  which they correspond.
This is equivalent to writing that two neurons $i$ and $j$ are  not  directly
connected  if and only if  $X_i$  and $X_j$  are  conditionally independent
with  respect to all  other variables. We remind  that two random  variables
$X_i$  and  $X_j$  are  said  to  be  conditionally independent to a (possibly
empty)  set of controlling random variables $Z$, denoted $X_i \perp X_j | Z$,
if the joint conditional probability distribution   $P_{X_i,  X_j   |   Z}$
factorizes  into   $P_{X_i|Z} P_{X_j|Z}$.  In particular,  when  $X_i$ and
$X_j$ are  conditionally independent  with respect  to all  the other
variables, denoted  $X_i \perp  X_j  | X^{-i,j}$,  then  the  observation  of
$X_i$  given  the knowledge  of  $X^{-i,j}$ provides  no  information  about
$X_j$,  and vice-versa.
